Precise spiking motifs in neurobiological and neuromorphic data

### 2022-11-08 manuscript.tex
Update on Overleaf.

\note{redo a figure similar to the thorpe code}%\includegraphics{images/roc.jpg}
\caption{Latency coding. An input analog profile is encoded in latencies: the higher the contrast, the shorter the latency. In this example, one generates at most one spike per neuron.
}\label{fig:roc}
In fact, similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{celebrini_dynamics_1993}. First-spike latency codes are a feasible mechanism for information transfer, even when biologically plausible estimates of stimulus onset are considered, for instance for sound localization~\citep{chase_first-spike_2007}.  Note also that timing is not entirely sensory or internal, but can be used as a general coding principle. In~\citep{safaie_turning_2020} for instance, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines. Timing, at least in animals, may thus be fundamentally embodied and situated.''  Many models have used these properties in temporal coding to build fast image categorization networks~\citep{gautrais_rate_1998,delorme_spikenet_1999, perrinet_coding_2004}. These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization~\citep{delorme_ultra-rapid_2000}. One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron by using the rank of neurons' activation~\citep{delorme_spikenet_1999,bonilla_analyzing_2022}. Another type of SNN using precise spike timing attempts to determine the structure of the network in order to minimize a cost function. This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extended in novel gradient-based methods. The surrogate gradient method is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}. This type of modeling often uses the classical architecture of image categorization developed in deep learning, while adapting it to the specificity of the event-based representation~\citep{goltz_fast_2021}. However, the performance of SNNs is still lagging that of analog networks and the question of the advantage of using spikes in machine learning and computer vision remains open. 
Achieving to efficiently implement this type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Retinas which represents the visual information as events, similarly to the representations described above and instead of using a basic frame-based representation~\citep{lichtsteiner_128-128_2008} (for a review, see~\citep{rasetto_challenges_2022}, this will be developped in Section~\ref{sec:neuromorphic}). % This work has been extended to include unsupervised learning capabilities, and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}. 
To alleviate this gap in computational neuroscience, we will further review the role of  precise spike timing in neural assemblies from a neuroscientific perspective in the next section.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Role of precise spike timing in neural assemblies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{One first hypothesis: synchronous firing in cell assemblies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
In his book ``Corticonics'',~\citet{abeles_corticonics_1991} queried whether the role of cortical neurons is to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns. The book gradually leads the reader from the macroscopic cortical anatomy and standard electro-physiological properties of single neurons to neural network models. While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code that we have already explored at the cellular level~\citep{abeles_role_1982,paugam-moisy_computing_2012}. The book then demonstrates that this could take the form of ``synfire chains,'' that is, the emergence of synchronous activity on subsets of neurons which could be propagated in a stable fashion. Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{harris_organization_2003}. One possible function of this synchronization may serve the binding of information distributed in the brain~\citep{singer_visual_1995, roelfsema_visuomotor_1997}.

Some experimental results show the emergence of synchrony, for instance in motor cortical function~\citep{riehle_spike_1997}. Interestingly, the authors showed that ``Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events. Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent. These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to movement direction and reaction time~\citep{grammont_spike_2003}. It is important to note that synchronous events tend to lock to LFP beta waves~\citep{denker_lfp_2018}, and were extended to larger assemblies~\citep{torre_synchronous_2016, torre_asset_2016} using statistical methods (see Section~\ref{sec:detection} for further details). Note also that synchronicity may explain some unintuitive results. Indeed, it has been shown that thalamo-cortical synapses are relatively weak compared to the amount of intra-cortical activity. However, this pathway is sufficient to drive the cortex, as this input is more often synchronously active~\citep{bruno_cortex_2006}.

In particular, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001} by considering the dynamics of leaky integrate-and-fire (LIF) neurons in different groups of similar size. Each neuron of one group is connected by an excitatory synapse to the next. When a pulse is elicited in the first group, this may generate a spike in the next group. Depending on the concentration of synaptic weight values, this new activity may get more or less synchronized than the previous pulse (as measured by the standard deviation of spike times in the pulse). Recursively applying this to a sequence of groups generates either a synfire propagation or not. A simple simulation of synfire propagation is shown in Figure~\ref{fig:diesman}. A crucial aspect of this emergence is in particular the dynamics of the spiking neuron model~\citep{gerstner_time_1995} and in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}. This balance was for instance modeled by feed-forward inhibition, a fine-scaled latency mechanism that is an essential ingredient in modelling push-pull effects in the primary visual cortex~\citep{kremkow_push-pull_2016}. Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003}, while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}. In particular, this was implemented as a computational neuroscience benchmark model using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013}. % (see Figure~\ref{fig:pynn}).
%
%----------------------------%
%\begin{figure}
%\centering
%\includegraphics[width=.7\textwidth]{figures/fnins-07-00011-g004.jpg}
%\caption{A pyNN implementation of a Synfire chain with feedforward inhibition. The background noise  is implemented as random Gaussian current. The state space gives %the duration of synfire chains as a function of the protocol's parameters.}\label{fig:pynn}
%\end{figure}
%----------------------------%
%----------------------------%
\begin{figure}
\centering
\includegraphics[width=.4\textwidth]{figures/Diesmann_et_al_1999.png}
\caption{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}. The model consists of 10 groups of 100 neurons each. A pulse with a given jitter is generated in the first group, which here generates a pulse after a certain processing delay in the second group with a lower jitter. This allows the propagation of the synchronous activity along the chain of the neural groups.}\label{fig:diesman}
\end{figure}
%----------------------------%

@@

As we saw, most SNN and in particular those adapted from analogous deep-learning like architectures, rely on an encoding of information based on a continuously varying firing-rate. A notable example of a SNN using precise spike timings is the polychronization model of~\citet{izhikevich_polychronization_2006}. This model is based on a random recurrent model of spiking neurons including synaptic delays chosen randomly from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}. It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}. It is worth mentioning that only the weights are changed using the STDP rule and that the set of delays is set randomly at initialization and that it is then frozen for the rest of the simulation. Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked but not synchronous firing patterns. The neurons composing a group discharge at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time. This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:PG}). According to the STDP rule, the neurons involved in this activity will see their synaptic weight increase and, thus, constitute a polychronous group. Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system. In other neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}. This representation has potentially a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs~\citep{izhikevich_polychronous_2009}.

@@

An extension of~\citep{izhikevich_polychronization_2006} was made in a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}. Indeed, while the original paper contained material within the text to reproduce the whole model (using MATLAB), it was not complete for the reproduction of all results presented in that manuscript. This more recent work details how this code was slightly corrected. It comes with a Python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process. Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}. Based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PG activation. They demonstrate the use of this method by investigating the claim that PGs might provide the foundation of a representational system. Stimulation of the trained network produces the activation of a PG, i.e. the propagation of firing activity through multiple layers due to convergent patterns of firing.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Learning synaptic delays}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Training delays ~\citep{state_training_2019}
A whole range of models of latency motifs detection are using STDP-based learning rules. For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition;~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning. In particular, to address how transmission delays and STDP can jointly determine these emergent pairwise activity-connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}. Moreover, modified STDP rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001}. They were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}. In particular, this has been applied for recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity. More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission. Here, we show how the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision. This demonstrates how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment~\citep{burkitt_predictive_2021}.

Specifically, a recent work proposed a bio-plausible unsupervised delay learning for extracting temporal features in spiking neural networks~\citep{nadafian_bio-plausible_2020}. The authors provided some mathematical proofs to show that our learning rule gives a neuron the ability to learn repeating spatio-temporal patterns. Furthermore, the experimental results of applying an STDP-based spiking neural network was applied on Random Dot Kinematogram. Another model of synaptic delay-weight plasticity~\citep{zhang_supervised_2020} integrates synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes. 
This was also presented by~\citep{wang_delay_2019} which propose a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance. Other models like that of~\citep{hazan_memory_2022} propose a weightless spiking neural networks that can perform a simple classification task which is applied to MNIST.
% Uses synthetic data to extend the existing Remote Supervised Method (ReSuMe) method.
In a recent paper~\citep{luo_supervised_2022}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of a single spiking neuron. In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.
In yet another computational model,~\citet{sun_learning_2016} show that the frequently activated polychronous neural groups can be learned efficiently by readout neurons with joint weight-delay spike-timing-dependent plasticity.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Real-world applications}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Most of these theoretical models are trying to reproduce neurobiological observations, while others propose applications to machine learning methods such as image processing. For instance, in a recent work~\citet{ghosh_spatiotemporal_2019} propose a two-stage unsupervised-supervised system for the categorization of spatio-temporal actions from an event-based stream. The first stage learns spatio-temporal convolutional filters targeted to minimize event-removal related changes to a local spatio-temporal spike-event pattern. The second stage takes the output of the spatio-temporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatio-temporal activity. For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work. Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters thus showing the potential of such representation.

%There are more applications to image processing using sparse spiking representations. Using the core computational unit defined, extension of the computation to a topographic representation similar to that observed in the primary visual cortex of mammals. design of micro-circuits with specific lateral interactions will allow us to design efficient micro-circuits for the sparse representation of images.
A predictive model defined in the framework of the free-energy principle was able to account for temporal delays in the nervous system, both at the sensory and motor levels~\citep{perrinet_active_2014}.  %
%and behavioral~\citep{khoei_flash-lag_2017} levels. Extending this knowledge to the optimization of delays in a SNN will provide a breakthrough in the efficiency of these networks.
Such a model was implemented at the network level by showing that motion-based prediction explains the role of tracking in motion extrapolation interpolation~\citep{khoei_motion-based_2013}. Note that this motion-based prediction is implemented in the weights of the neural network by defining connections between neurons with specific delays.  Two neurons which were selective to specific motions were connected if the delay was coherent with the change in the position of their respective receptive fields~\citep{hogendoorn_predictive_2019}. This was also implemented in a spiking neural network and reproduced the observation that neural activity was maintained during the trajectory of a smoothly moving dot even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}. This led in particular to the proposal that such delay-based computations could explain diverse perceptual mechanisms, such as the flash-lag illusion~\citep{khoei_flash-lag_2017}. 
In the following section, we will review some more applications of using precise spiking motifs in neuromorphic engineering, but will first study current evidence in neurobiology, that is, from actual recordings in biological tissues.

@@

>>>>

%----------------------------%
\begin{figure}
\centering
\note{redo a figure similar to the thorpe code}%\includegraphics{images/roc.jpg}
\caption{Latency coding. An input analog profile is encoded in latencies: the higher the contrast, the shorter the latency. In this example, one generates at most one spike per neuron.
}\label{fig:roc}
\end{figure}
%----------------------------%
In fact, similar results have been demonstrated through neurophysiological recordings in the primary visual cortex and show that different levels of visual activity will induce different levels of neuronal discharge latency in the primary visual area~\citep{celebrini_dynamics_1993}. First-spike latency codes are a feasible mechanism for information transfer, even when biologically plausible estimates of stimulus onset are considered, for instance for sound localization~\citep{chase_first-spike_2007}.  Note also that timing is not entirely sensory or internal, but can be used as a general coding principle. In~\citep{safaie_turning_2020} for instance, they found that ``timing accuracy was improved when the environment afforded cues that rats can incorporate into motor routines. Timing, at least in animals, may thus be fundamentally embodied and situated.''  Many models have used these properties in temporal coding to build fast image categorization networks~\citep{gautrais_rate_1998,delorme_spikenet_1999, perrinet_coding_2004}. These models take the form of artificial spiking neural networks (SNNs) and have been able to demonstrate their practical applications for image categorization~\citep{delorme_ultra-rapid_2000}. One of these is the SpikeNet algorithm, which uses a purely temporal approach by encoding information using one spike per neuron by using the rank of neurons' activation~\citep{delorme_spikenet_1999,bonilla_analyzing_2022}. Another type of SNN using precise spike timing attempts to determine the structure of the network in order to minimize a cost function. This was implemented in the SpikeProp algorithm~\citep{bohte_error-backpropagation_2002} and has been extended in novel gradient-based methods. The surrogate gradient method is now widely used in methods that attempt to transfer performance from CNNs to SNNs~\citep{zenke_remarkable_2021}. This type of modeling often uses the classical architecture of image categorization developed in deep learning, while adapting it to the specificity of the event-based representation~\citep{goltz_fast_2021}. For instance,~\citep{kheradpisheh_stdp-based_2018} implements a STDP-based spiking deep convolutional neural networks for object recognition or~\citep{tavanaei_representation_2018} develops a form of spike-based, competitive learning applied for unsupervised learning.

However, the performance of SNNs is still lagging compared to that of analog networks and the question of the advantage of using spikes in machine learning and computer vision remains open. Achieving to efficiently implement this type of modeling is extremely important with respect to the development of a new generation of cameras called Silicon Retinas which represents the visual information as events, similarly to the representations described above and instead of using a basic frame-based representation~\citep{lichtsteiner_128-128_2008} (for a review, see~\citep{rasetto_challenges_2022}, this will be developped in Section~\ref{sec:neuromorphic}). % This work has been extended to include unsupervised learning capabilities, and we have recently developed a SNN architecture that allows to categorize images of different classes in only a few spikes~\citep{grimaldi_homeostatic_2021,grimaldi_robust_2022}. 
To alleviate this gap in computational neuroscience, we will further review the role of  precise spike timing in neural assemblies from a neuroscientific perspective in the next section.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Role of precise spike timing in neural assemblies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{One first hypothesis: synchronous firing in cell assemblies}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
In his book ``Corticonics'',~\citet{abeles_corticonics_1991} queried whether the role of cortical neurons is to integrate synaptic inputs or rather to detect coincidences in temporal spiking patterns. The book gradually leads the reader from the macroscopic cortical anatomy and standard electro-physiological properties of single neurons to neural network models. While the first hypothesis favors the rate coding theory, the second possibility highlights the need for temporal precision in the neural code that we have already explored at the cellular level~\citep{abeles_role_1982,paugam-moisy_computing_2012}. The book then demonstrates that this could take the form of ``synfire chains,'' that is, the emergence of synchronous activity on subsets of neurons which could be propagated in a stable fashion. Since this date, multiple experimental observations have suggested the existence of this precise zero-phase-lag spike synchronization in a defined subset of neurons~\citep{harris_organization_2003}. One possible function of this synchronization may serve the binding of information distributed in the brain~\citep{singer_visual_1995, roelfsema_visuomotor_1997}.

Some experimental results show the emergence of synchrony, for instance in motor cortical function~\citep{riehle_spike_1997}. Interestingly, the authors showed that ``Accurate spike synchronization occurred in relation to external events (stimuli, movements) and was commonly accompanied by discharge rate modulations but without precise time locking of the spikes to these external events. Spike synchronization also occurred in relation to purely internal events (stimulus expectancy), where firing rate modulations were distinctly absent. These findings indicate that internally generated synchronization of individual spike discharges may subserve the cortical organization of cognitive motor processes.'' Moreover, such emergence could change over the learning period involved in learning a task~\citep{kilavik_long-term_2009} and showed some tuning to movement direction and reaction time~\citep{grammont_spike_2003}. It is important to note that synchronous events tend to lock to spatio-temporal patterns of neural activity called \href{https://en.wikipedia.org/wiki/Local_field_potential}{LFP} beta waves~\citep{denker_lfp_2018}, and were extended to larger assemblies~\citep{torre_synchronous_2016, torre_asset_2016} using statistical methods (see Section~\ref{sec:detection} for further details). Note also that synchronicity may explain some unintuitive results. Indeed, it has been shown that thalamo-cortical synapses are relatively weak compared to the amount of intra-cortical activity. However, this pathway is sufficient to drive the cortex, as this input is more often synchronously active~\citep{bruno_cortex_2006}.

In particular, it was shown that a simple model may allow the propagation of such synfire chains~\citep{gewaltig_propagation_2001} by considering the dynamics of leaky integrate-and-fire (LIF) neurons in different groups of similar size. Each neuron of one group is connected by an excitatory synapse to the next. When a pulse is elicited in the first group, this may generate a spike in the next group. Depending on the concentration of synaptic weight values, this new activity may get more or less synchronized than the previous pulse (as measured by the standard deviation of spike times in the pulse). Recursively applying this to a sequence of groups generates either a synfire propagation or not. A simple simulation of synfire propagation is shown in Figure~\ref{fig:diesman}. A crucial aspect of this emergence is in particular the dynamics of the spiking neuron model~\citep{gerstner_time_1995} and in particular the balance between excitation and inhibition~\citep{azouz_stimulus-selective_2008}. This balance was for instance modeled by feed-forward inhibition, a fine-scaled latency mechanism that is an essential ingredient in modelling so-called push-pull effects in the primary visual cortex~\citep{kremkow_push-pull_2016}. Further models have shown that such synfire chains could be embedded in topographies~\citep{aviel_embedding_2003}, while others used conductance-based neurons with feed-forward inhibition to improve the robustness of the propagation~\citep{kremkow_functional_2010}. In particular, this was implemented as a computational neuroscience benchmark model using the pyNN language~\citep{davison_pynn_2008} both in CPU-based and neuromorphic hardware~\citep{pfeil_six_2013}. % (see Figure~\ref{fig:pynn}).
%
%----------------------------%
%\begin{figure}
%\centering
%\includegraphics[width=.7\textwidth]{figures/fnins-07-00011-g004.jpg}
%\caption{A pyNN implementation of a Synfire chain with feedforward inhibition. The background noise  is implemented as random Gaussian current. The state space gives %the duration of synfire chains as a function of the protocol's parameters.}\label{fig:pynn}
%\end{figure}
%----------------------------%
%----------------------------%
\begin{figure}
\centering
\includegraphics[width=.4\textwidth]{figures/Diesmann_et_al_1999.png}
\caption{Simulation of a synfire propagation \href{https://brian2.readthedocs.io/en/stable/examples/frompapers.Diesmann_et_al_1999.html}{using Brian}. The model consists of 10 groups of 100 neurons each. A pulse with a given jitter is generated in the first group, which here generates a pulse after a certain processing delay in the second group with a lower jitter. This allows the propagation of the synchronous activity along the chain of the neural groups.}\label{fig:diesman}
\end{figure}
%----------------------------%

@@

As we saw, most SNN and in particular those adapted from analogous deep-learning like architectures, rely on an encoding of information based on a continuously varying firing-rate. A notable example of a SNN using precise spike timings is the polychronization model of~\citet{izhikevich_polychronization_2006}. This model is based on a random recurrent model of spiking neurons including synaptic delays chosen randomly from a range of biologically realistic delays (from $0$ to $20~\ms$) and whose weights evolved with a Spike-Time Dependent Plasticity (STDP) learning rule~\citep{markram_regulation_1997}. It is consensus that spike timing (STDP) plays a crucial role in the development of synaptic efficacy for many different kinds of neurons~\citep{caporale_spike_2008}. It is worth mentioning that only the weights are changed using the STDP rule and that the set of delays is set randomly at initialization and that it is then frozen for the rest of the simulation. Delays are defined as the total time taken for a spike to be conducted from one presynaptic neuron's soma to the efferent postsynaptic neuron's soma. Due to the interplay between the delays and STDP, the spiking neurons spontaneously self-organize into groups and generate patterns of stereotypical polychronous activity, i.e.~exhibit reproducible time-locked but not synchronous firing patterns. The neurons composing a group discharge at different times, but due to the heterogeneous delays, the spikes reach the postsynaptic neuron at the same time. This synchrony leads to the summation of the excitatory post-synaptic potentials evoked by each spike, and thus to the crossing of the voltage threshold and to the discharge of a spike (see Figure~\ref{fig:PG}). According to the STDP rule, the neurons involved in this activity will see their synaptic weight increase and, thus, constitute a polychronous group. Interestingly, thanks to the fact that a neuron can be involved in different polychronous groups, the number of coexisting polychronous groups far exceeds the number of neurons in the network, resulting in an unprecedented memory capacity of the system. In other neuronal models, an efficient use or detection of these spatio-temporal patterns embedded in the spike train comes with the integration of heterogeneous delays~\citep{guise_bayesian_2014,zhang_supervised_2020}. This representation has potentially a much greater information capacity in comparison to other neural coding approaches through their connectivity and the possible coexistence of numerous superposed PGs~\citep{izhikevich_polychronous_2009}.

@@

An extension of~\citep{izhikevich_polychronization_2006} was made in a very detailed work aiming at reproducing the polychronization model~\citep{pauli_reproducing_2018}. Indeed, while the original paper contained material within the text to reproduce the whole model (using MATLAB), it was not complete for the reproduction of all results presented in that manuscript. This more recent work details how this code was slightly corrected. It comes with a Python code and a version control system detailing the whole process used to give provenance to the different steps in this scientific process. Another recent work gives a Bayesian account in a similar model~\citep{guise_bayesian_2014}. Based on the fact that previous methods for studying polychronous groups' activation response to stimuli have been limited by the template-based methods used to identify PG activation, the authors outline a new method that overcomes these difficulties by establishing a probabilistic interpretation of PG activation. They demonstrate the use of this method by investigating the claim that PGs might provide the foundation of a representational system. Stimulation of the trained network produces the activation of a PG, i.e. the propagation of firing activity through multiple layers due to convergent patterns of firing. While extending the original method, these methods reveal shortcomings that we will try to analyze in the rest of this section.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Learning synaptic delays}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Training delays ~\citep{state_training_2019}
First, the original model use a simple STDP rule while a whole range of STDP-based learning rules may implement precise spiking motifs detection. For instance, to address how transmission delays and STDP can jointly determine these emergent pairwise activity-connectivity patterns, a recent study analyzed phase synchronization properties and coupling symmetry between two bidirectionally coupled neurons using both phase oscillator and conductance-based neuron models~\citep{madadi_asl_delay-dependent_2022}. Moreover, modified STDP rules have been used for synchronous coherence detection~\citep{perrinet_coherence_2002}, for the learning of specific receptive fields~\citep{perrinet_networks_2001}. They were also extended to recurrent Neuronal Networks~\citep{gilson_stdp_2010} or delay selection~\citep{datadien_right_2011}. In particular, this has been applied for recurrent networks of spiking neurons receiving oscillatory inputs~\citep{kerr_delay_2013} which targets for the selective potentiation of recurrent connections with different axonal and dendritic delays during oscillatory activity. More generally, our ability to track and respond to rapidly changing visual stimuli, such as a fast-moving tennis ball, indicates that the brain is capable of extrapolating the trajectory of a moving object to predict its current position, despite the delays that result from neural transmission. Specifically, the neural circuits underlying this ability can be learned through spike-timing-dependent synaptic plasticity and that these circuits emerge spontaneously and without supervision, demonstrating how the neural transmission delays can, in part, be compensated to implement the extrapolation mechanisms required to predict where a moving object is at the present moment~\citep{burkitt_predictive_2021}.

Furthermore, a recent work proposed a bio-plausible unsupervised delay learning for extracting temporal features in spiking neural networks~\citep{nadafian_bio-plausible_2020}. The authors provided some mathematical proofs to show that their learning rule gives the ability to learn repeating spatio-temporal patterns. Applying this STDP-based rule on delays to the spiking neural network, experimental results were validated on a simple motion detection task. Another model of synaptic delay-weight plasticity~\citep{zhang_supervised_2020} integrates synaptic delay plasticity into supervised learning and proposes a novel learning method that adjusts both the synaptic delays and weights of the learning neurons to make them fire precisely timed spikes. This was also presented by~\citep{wang_delay_2019} which propose a supervised delay learning algorithm for spiking neurons with temporal encoding, in which both the weight and delay of a synaptic connection can be adjusted to enhance the learning performance. Other models like that of~\citep{hazan_memory_2022} propose a weightless spiking neural networks that can perform a simple classification task which is applied to MNIST. In a recent paper~\citep{luo_supervised_2022}, authors propose a gradient descent-based learning algorithm for synaptic delays to enhance the sequential learning performance of a single spiking neuron. In this algorithm, information is encoded in the relative timing of individual neuronal spikes, and learning is performed based on the exact derivatives of the postsynaptic spike times with respect to presynaptic spike times.
In yet another computational model,~\citet{sun_learning_2016} show that the frequently activated polychronous neural groups can be learned efficiently by readout neurons with joint weight-delay spike-timing-dependent plasticity.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Real-world applications}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
A second shortcoming of models derived from the polychronization model is their lack of applications in real-world scenarios. Indeed, most of these theoretical models are trying to reproduce neurobiological observations, while applications to machine learning methods such as image processing would further prove their plausibility. For instance, in a recent work~\citet{ghosh_spatiotemporal_2019} propose a two-stage unsupervised-supervised system for the categorization of spatio-temporal actions from an event-based stream. The first stage learns spatio-temporal convolutional filters targeted to minimize event-removal related changes to a local spatio-temporal spike-event pattern. The second stage takes the output of the spatio-temporal filters as an input example containing multiple feature channels, and proceeds to train a classifier for recognition of spatio-temporal activity. For testing the system, two datasets are considered: DVS Gesture and a new action recognition dataset recorded for this work. Results demonstrate the ability of the system to outperform the state-of-the-art in event-based gesture recognition, along with demonstrating superior performance to other alternative ways of obtaining the first stage filters, thus showing the potential of such representation.

There are more applications to image processing using spiking neural networks based on the design of micro-circuits with specific lateral interactions embedded with spatially anisotropic connections. Using this core computational unit defined, and extending to computations on a topographic representation similar to that observed in the primary visual cortex of mammals, the anisotropic rules implemented a form of delayed activation. This result was based on a predictive model defined in the Bayesian framework (the so-called free-energy principle) which was able to account for temporal delays in the  system, both at the sensory and motor levels~\citep{perrinet_active_2014}, and in particular that ``the application of delay operators just means changing synaptic connection strengths to take different mixtures of generalized sensations and their prediction errors.'' Such a model was implemented at the network level and applied to various motion detection tasks. Two neurons which were selective to specific motions were connected if the delay was coherent with the change in the position of their respective receptive fields~\citep{hogendoorn_predictive_2019}.  This was also implemented in a neural mass model, showing that such anisotropic connectivity may explain the emergence of tracking~\citep{khoei_motion-based_2013}, and further explored in a spiking neural network which reproduced the observation that neural activity was maintained during the trajectory of a smoothly moving dot even if it was momentarily blanked~\citep{kaplan_anisotropic_2013}. This led in particular to the proposal that such delay-based computations could explain diverse perceptual mechanisms, such as the so-called flash-lag illusion~\citep{khoei_flash-lag_2017}. However, these latter models used parametric rules for defining the wrights. Extending such mechanisms with the ability of learning delays in a SNN will provide a breakthrough in the efficiency of these networks, and we will explore results from neurobiology to get better insights on that aspect.

@@

In the following section, we will first review current evidence of taking advantage of spiking motifs in neurobiology, that is, from actual recordings in biological tissues. % some more applications of using precise spiking motifs in neuromorphic engineering, but will first study
